Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2017-01-23 body/abstract.tex
andrea s last modif plus 250 words abstract

We tackle this issue with our novel  idea, the reachability condition. This condition allows us to define
a geometric approximation of the contact manifold, which is of low dimension and can be sampled efficiently.
Informally, the condition verifies that the root configuration of a robot is close, but not too close to obstacles: close to allow contact creation, not too close to avoid collision. Thanks to this condition we decompose the hard contact planning problem into two
simpler sub-problems: first, planning a guide path for the robot root without considering the whole-body configuration, using a sampling-based algorithm; then, generating a discrete sequence of whole-body configurations in static equilibrium along this path, using a new deterministic
contact-selection algorithm.

Our approach results from the pragmatic choice of favoring efficiency over completeness, which we justify empirically: in a
few seconds, with satisfying success rates, we generate complex contact plans for various scenarios and robots, namely
HRP-2, HyQ, and a dexterous hand. These plans are then validated either in dynamic simulations, or on the real HRP-2 robot, thus demonstrating 
the first interactive multi-contact planner.

>>>>

We tackle this issue by introducing the reachability condition. This condition allows us to define
a geometric approximation of the contact manifold, which is of low dimension and can be sampled efficiently.
Informally, the condition verifies that the root configuration of a robot is close, but not too close to obstacles: close to allow contact creation, not too close to avoid collision. Thanks to this condition we decompose the hard contact planning problem into two
sub-problems: first, planning a guide path for the root without considering the whole-body configuration, using a sampling-based algorithm; then, generating a discrete sequence of whole-body configurations in static equilibrium along this path, using a deterministic contact-selection algorithm.

Our approach results from the pragmatic choice of favoring efficiency over completeness, which we justify empirically: in a
few seconds, with high success rates, we generate complex contact plans for various scenarios and robots: HRP-2, HyQ, and a dexterous hand. These plans are then validated either in dynamic simulations, or on the real HRP-2 robot, thus demonstrating 
the first interactive multi-contact planner.

### 2017-01-23 body/appendix_optim.tex
andrea s last modif plus 250 words abstract

% !TEX root =  ../main.tex
\section{Addressing $\mathcal{P}_3$}
\label{app:optim}
The stair climbing and the standing up scenarios were validated with the trajectory optimization scheme provided in \citeauthor{Carpentier2016}. 
\deladp{The optimization is formulated as an optimal control problem that aims at generating a valid trajectory for the center of mass.}
To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox}), which can be integrated directly in our motion planner software. This formulation uses the center of mass acceleration and angular momentum as input variables, while previously the contact forces were used.
The center-of-mass trajectory resulting from the optimization is then turned into a collision-free whole-body trajectory.

>>>>

% !TEX root =  ../main.tex
\section{Addressing $\mathcal{P}_3$}
\label{app:optim}
The stair climbing and the standing up scenarios were validated with the trajectory optimization scheme provided in \citeauthor{Carpentier2016}. 
%~ \deladp{The optimization is formulated as an optimal control problem that aims at generating a valid trajectory for the center of mass.}
To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox}), which can be integrated directly in our motion planner software. This formulation uses the center of mass acceleration and angular momentum as input variables, while previously the contact forces were used.
The center-of-mass trajectory resulting from the optimization is then turned into a collision-free whole-body trajectory.

### 2017-01-23 body/conclusion.tex
andrea s last modif plus 250 words abstract

The main source of computational cost identified in previous works is the verification of \equilibriumfeasibility. 
The strong assumption of our work is that for the class of problems we consider \contactreachability\ implies \equilibriumfeasibility.
Our scenarios show that the  assumption is verified in the majority of cases when at least one contact surface is \textit{quasi flat}~\cite{Prete2016}, that is when
the friction cone of the contact surface contains the direction opposite to the gravity.
Figure~\ref{fig:dedefeas} illustrates this observation, demonstrated empirically by the third column of Table~\ref{tab:requestpercent}. In the worst case, in our experiment
the assumption was verified for 82\% of the total amount of trials.

In the example of \cite{Bouyarmane2009}, the verification of \equilibriumfeasibility\ implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$, requiring
several minutes of planning. Our method, in comparison, takes from a few milliseconds to several seconds.

These results clearly justify the pragmatic decisions of this paper. What is lost in completeness is largely compensated by the computation times we obtain.

>>>>

Another source of computational cost identified in previous works is the verification of \equilibriumfeasibility. 
The strong assumption of our work is that for the class of problems we consider \contactreachability\ implies \equilibriumfeasibility.
Our scenarios show that the  assumption is verified in the majority of cases when at least one contact surface is \textit{quasi flat}~\cite{Prete2016}, that is when
the friction cone of the contact surface contains the direction opposite to the gravity.
Figure~\ref{fig:dedefeas} illustrates this observation, demonstrated empirically by the third column of Table~\ref{tab:requestpercent}. In the worst case, in our experiment
the assumption was verified for 82\% of the total amount of trials that verified \contactreachability\.
In the example of \cite{Bouyarmane2009}, the verification of \equilibriumfeasibility\ implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$, requiring
several minutes of planning. Our method, in comparison, takes from a few milliseconds to several seconds.

These results clearly justify the pragmatic decisions of this paper.

### 2017-01-23 body/contact.tex
andrea s last modif plus 250 words abstract

\subsection{Definition of a contact sequence}
In previous contributions\adnote{Add reference here}, a contact plan is defined as a sequence of quasi-static equilibrium configurations
for each contact phase. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact phases. Furthermore it removes the need to have single-support quasi-static configurations as in the example. 
As shown in the companion video, this allows our framework to produce dynamic motions.

>>>>

\subsection{Definition of a contact sequence}
In previous contributions~\citep{DBLP:conf/iser/EscandeKMG08}, a contact plan is defined as a sequence of quasi-static equilibrium configurations
for each contact phase. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact phases. Furthermore it removes the need to have single-support quasi-static configurations as in the example. 
As shown in the companion video, this allows our framework to produce dynamic motions.
